Windowed Developer Topic Analysis

### 2009-04-07 lda-paper.tex
Latest changes

Our contributions include:
\begin{itemize}
\item Demonstrating the value of windowed topic analysis
\item Multiple visualizations of trends over time
\end{itemize}

@@

Some terms we will be using for the rest of the paper include:
message, word distribution, topic and trend. \emph{Messages} are
blocks of text written by developers. In this paper messages will be
the CVS commit comments made when the user commits changes to files in
a repository. Word distributions are the summary of a message by word
count. The word distribution will be over all messages
looked at but since most words will not appear in a message, a word
distribution is effectively a word count divided by the message
size. A topic is a unique word distribution, but effectively a set of
words unique to an independent behavior in the text corpus, in this
paper we often summarize topics by their top 10 more frequent words.
A trend is a one or more similar topics which reoccur over
time. Trends are important because they are sets of topics which
reoccur over the development of a project.

@@

Latent Dirichlet Allocation (LDA)~\cite{944937} is a competitor to
Latent Semantic Indexing (LSI), probabilistic Latent Semantic Indexing
(pLSI) and semantic clustering. It is used for document modelling,
document clustering and collaborative filtering. What LDA attempts to
do is model a documents as mixture model of topics.  The common use of
LDA in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
is to extract topic clusters from documents such as methods, bug
reports, source code and files.

Each of these, LDA and LSI, extract topic clusters
from the documents that are independent from one another. LDA doesn't
name these clusters but the clusters consist of words whose
relationship is often obvious to someone familiar with the
corpus. We could swap LDA for LSI as produce similar results. Our data is posed as
documents with word distributions (word counts per documents) and LDA
extracts distinct topics (clusters of words) from the documents.

@@

Linstead et al has proposed an author source code model using
LDA~\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}. They used an
author-topic (AT) model~\cite{1036902} to infer the association or
expertise of a particular developer with a particular code topic
extracted from the source control system.

Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help query for
finding related documents during bug localization tasks. They used LDA
to build a topic model and then query against it using sample
documents, the query would the proportion of topics that were
associated with the query and thus the related documents.

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
This is when concepts are automatically extracted from source code or
documents are related to part of the source code.  Concept location is
how high level concepts relate to low level entities like source code,
for instance when fixing a bug how do the concepts in the bug report
map back to the source code?  Semantic Clustering has also been used
for similar reason~\cite{1698774,1566153}.

>>>>

Our contributions include:
\begin{itemize}
\item Demonstrating the value of windowed topic analysis using trends
\item Multiple visualizations topics and trends over time
\end{itemize}

@@

Some terms we will be using for the rest of the paper include:
message, word distribution, topics and trends. \emph{Messages} are
blocks of text written by developers. In this paper messages will be
the CVS commit comments made when the user commits changes to files in
a repository. \emph{Word distributions} are the summary of a message
by word count. The word distribution will be over all messages looked
at but since most words will not appear in a message, a word
distribution is effectively a word count divided by the message
size. A \emph{topic} is essentially a set of words which form a word
distribution that is unique and independant within the set of
documents in our total corpus. One could think of a topic as the
center distribution of a group of messages. In this paper we often
summarize topics by their top 10 most frequent words in their word
distribution.  A \emph{trend} is a one or more similar topics which
reoccur over time. Trends are important because they are sets of
topics which reoccur over the development of a project.

@@

In terms of clustering and finding topic distributions, Latent
Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
Indexing (LSI), probabilistic Latent Semantic Indexing (pLSI) and
semantic clustering. These tools is used for document modelling,
document clustering and collaborative filtering. What LDA attempts to
do is encode documents as mixture model of topics.  LDA is used in
software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
is to extract topic clusters from documents such as methods, bug
reports, source code and files.

Each of these, LDA, LSI and semantic clustering, extract topic
clusters from the documents that are independent from one another. LDA
doesn't name these clusters but the clusters consist of words whose
relationship is often obvious to someone familiar with the corpus. We
could swap LDA for LSI or semantic clustering and produce similar
results. Our data is posed as documents with word distributions (word
counts per documents) and LDA extracts distinct topics (clusters of
words) from the documents.

@@

In order to infer or associate the expertise of an author with topics
extracted from SCS, Linstead et al proposed an author-sourcecode model
using LDA~\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}. This model
is essentially an extension of the author-topic (AT)
model~\cite{1036902}, which associated authors and topics extracted
with LDA.

Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help bug
localization by querying for related documents to a bug's topic. They
used LDA to build a topic model and then queried against it using
sample documents. These queries would indicate the proportion of
topics that were associated with the query and thus possible the
related documents.

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
This is where concepts are automatically extracted from source code or
documents related to the source code.  Concept location is how high
level concepts relate to low level entities like source code, for
instance when fixing a bug how do the concepts in the bug report map
back to the source code?  Semantic Clustering has also been used for
similar reasons~\cite{1698774,1566153} as it is similar to LSI.
